Subject-level developmental relationships

### 2024-08-30 1_get_data.qmd
updates from more hacking

>>>>

Add CDIs.

This is very slightly broken due to some repeated fields. We use `values_fn = mean` to suppress them.

```{r}
load(here("..", "peekbank-method", "cached_intermediates", "1_cdi_subjects.Rds"))

d_cdi <- cdi_data |>
  select(administration_id, native_language, measure, language, 
         CDI_percent) |> # rawscore is form-dependent
  filter(language == "English (American)", native_language == "eng") |>
  pivot_wider(names_from = "measure", values_from = "CDI_percent", 
              values_fill = NA, values_fn = mean)
```

```{r}
d_sub <- d_sub |>
  left_join(d_cdi) |>
  mutate(log_age = log(age)) |>
  ungroup()
```

Print descriptives.

```{r}
print(paste(length(unique(d_sub$subject_id)), "children,",
            sum(!is.na(d_sub$prod)), "with CDI data"))
print(paste(length(unique(d_sub$administration_id)), "administrations"))
print(paste(length(unique(d_sub$dataset_name)), "datasets"))
```

### 2024-08-30 2_developmental_analysis.qmd
updates from more hacking

```{r}
d_trial |>
ggplot(aes(x = age, y = rt)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  scale_x_log10() +
  scale_y_log10()
```

@@

```{r}
d_sub |>
  pivot_longer(contains("accuracy"), 
               values_to = "accuracy", names_to = "measure") |>
ggplot(aes(x = age, y = accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~measure)
```

RT - but is it the mean of the log or the log of the mean? We choose log of the mean because this fits a log-normal distribution.

```{r}
d_sub |>
ggplot(aes(x = age, y = log_rt)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  scale_x_log10()  + 
  ylab("mean log RT")
```

@@

>>>>

```{r}
d_trial |>
  pivot_longer(c("rt","log_rt"), 
               values_to = "rt", names_to = "measure") |>
ggplot(aes(x = age, y = rt)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  scale_x_log10() +
  facet_wrap(~measure, scales = "free_y")
```
## Stats

```{r}
mods <- list(log_log = lm(log_rt ~ log_age, data = d_trial),
             log_lin = lm(log_rt ~ age, data = d_trial),
             lin_log = lm(rt ~ log_age, data = d_trial),
             lin_lin = lm(rt ~ age, data = d_trial))

models <- map_df(mods, ~broom::glance(.x)) |>
  mutate(model = names(mods))

kable(models, digits = 4)
```

@@

```{r}
d_sub |>
  pivot_longer(contains("accuracy"), 
               values_to = "accuracy", names_to = "measure") |>
ggplot(aes(x = age, y = accuracy)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() + 
  facet_wrap(~measure)
```
RTs

```{r}
d_sub |>
  pivot_longer(c("rt","log_rt"), 
               values_to = "rt", names_to = "measure") |>
ggplot(aes(x = age, y = rt)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  scale_x_log10()  + 
  ylab("mean log RT") + 
  facet_wrap(~measure, scales = "free_y")
```

@@

## Stats

```{r}
mods_sub <- list(log_log = lm(log_rt ~ log_age, data = d_sub),
             log_lin = lm(log_rt ~ age, data = d_sub),
             lin_log = lm(rt ~ log_age, data = d_sub),
             lin_lin = lm(rt ~ age, data = d_sub))

models_sub <- map_df(mods_sub, ~broom::glance(.x)) |>
  mutate(model = names(mods_sub))

kable(models_sub, digits = 4)
```

# Variability

```{r}
d_trial |>
  group_by(administration_id, age) |>
  summarise (sd_rt = sd(rt, na.rm=TRUE), 
             sd_acc = sd(long_window_accuracy, na.rm=TRUE)) |>
  pivot_longer(contains("sd"), 
               values_to = "sd", names_to = "measure") |>
ggplot(aes(x = age, y = sd)) + 
  geom_point(alpha = .05) + 
  geom_smooth() +
  geom_smooth(method = "lm", col = "green") +
  # geom_hline(yintercept = .5, lty = 2) + 
  scale_x_log10() +
  facet_wrap(~measure, scales = "free_y")
```

### 2024-08-30 3_developmental_relationships.qmd
updates from more hacking

```{r}
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(source(here("helper","common.R")))
d_sub <- readRDS(here("cached_intermediates","1_d_sub.Rds"))
load(here("..", "peekbank-method", "cached_intermediates", "1_cdi_subjects.Rds"))
```

This is very slightly broken due to some repeated fields. We use `values_fn = mean` to suppress them.

```{r}
d_cdi <- cdi_data |>
  select(administration_id, native_language, measure, language, 
         CDI_percent) |> # rawscore is form-dependent
  filter(language == "English (American)", native_language == "eng") |>
  pivot_wider(names_from = "measure", values_from = "CDI_percent", 
              values_fill = NA, values_fn = mean)
```

```{r}
d_sub_cdi <- d_sub |>
  left_join(d_cdi) |>
  mutate(log_age = log(age)) |>
  ungroup()
```

@@

```{r, fig.width=9, fig.height=9}
GGally::ggpairs(select(d_sub_cdi, age, log_age, rt, log_rt, 
                       short_window_accuracy, long_window_accuracy, 
                       short_window_elogit, long_window_elogit, prod, comp), 
                progress = FALSE, lower = list(continuous = GGally::wrap("points", alpha = 0.03)))

@@

```{r, fig.width=9, fig.height=9}
GGally::ggpairs(select(d_sub_cdi, log_age, rt,  
                       long_window_accuracy, short_window_accuracy, prod, comp), 
                progress = FALSE, lower = list(continuous = GGally::wrap("points", alpha = 0.03)))

@@

```{r}
m_rt <- mean(d_sub_cdi$rt, na.rm=TRUE)
sd_rt <- sd(d_sub_cdi$rt, na.rm=TRUE)
d_sub_cdi$factor_rt <- cut(d_sub_cdi$rt, 
                           breaks = quantile(d_sub_cdi$rt, 
                                             probs = c(0, .25, .5, .75, 1), 
                                             na.rm= TRUE), 
                           labels = c("0-25", "25-50", "50-75", "75-1"))

ggplot(filter(d_sub_cdi, !is.na(rt), !is.na(factor_rt)),
       aes(x = age, 
           y = long_window_accuracy, 
           col = factor_rt)) +
  geom_point(alpha = .3) +
  geom_smooth() +
  scale_x_log10() + 
  geom_hline(yintercept = .5, lty = 2)+
  viridis::scale_color_viridis(discrete = TRUE, 
                               name = "RT quantile") + 
  xlab("Age (log months)") + 
  ylab("Accuracy (long window)") + 
  theme(legend.position = "bottom")

@@

d_sub_cdi$age_group <- cut(d_sub_cdi$log_age, 
                           breaks = log(c(12,24,36,48,60)), 
                           labels = c("12-24","24-36","36-48","48-60"), 
                           include_lowest = TRUE)

ggplot(filter(d_sub_cdi, !is.na(age_group)), 
       aes(x = rt, 
           y = long_window_accuracy, 
           col = age_group)) +
  geom_point(alpha = .3) +
  geom_smooth() +
  geom_hline(yintercept = .5, lty = 2)+
  viridis::scale_color_viridis(discrete = TRUE, 
                               name = "age group")
```

@@

ggplot(filter(d_sub_cdi, 
              !is.na(rt), !is.na(factor_rt), 
              dataset_name %in% datasets_with_age_variation$dataset_name),
       aes(x = age, 
           y = long_window_accuracy, 
           col = factor_rt)) +
  geom_point(alpha = .3) +
  geom_smooth(se= FALSE, method = "lm") +
  scale_x_log10() + 
  geom_hline(yintercept = .5, lty = 2)+
  viridis::scale_color_viridis(discrete = TRUE, 
                               name = "RT quantile") +
  facet_wrap(~dataset_name) + 
  xlab("Age (log months)") + 
  ylab("Accuracy (long window)")

@@

```{r}
d_sub_cdi_mat <- d_sub_cdi |>
  ungroup() |>
  select(log_age, rt,  long_window_accuracy, prod)

d_sub_cdi_mat <- d_sub_cdi_mat |>
  filter(complete.cases(d_sub_cdi_mat)) |>
  as.matrix()

acc_rt_prc <- prcomp(x = d_sub_cdi_mat, scale = TRUE)

@@

```{r}
mod1 <- fa(d_sub_cdi_mat, 1)
mod2 <- fa(d_sub_cdi_mat, 2)
```

@@

```{r}
fa.parallel(d_sub_cdi_mat)
```

@@

```{r}
d_sub_cdi_mat_comp <- d_sub_cdi |>
  ungroup() |>
  select(log_age, rt,  long_window_accuracy, comp)

d_sub_cdi_mat_comp <- d_sub_cdi_mat_comp |>
  filter(complete.cases(d_sub_cdi_mat_comp)) |>
  as.matrix()

acc_rt_prc_comp <- prcomp(x = d_sub_cdi_mat_comp, scale = TRUE)

@@

```{r}
mod1_comp <- fa(d_sub_cdi_mat_comp, 1)
mod2_comp <- fa(d_sub_cdi_mat_comp, 2)
```

```{r}
fa.parallel(d_sub_cdi_mat_comp)
```

@@

```{r}
longitudinal <- d_sub_cdi |>
  group_by(dataset_name, subject_id) |>
  count() |>
  filter(n > 1)

d_sub_long <- filter(d_sub_cdi, subject_id %in% longitudinal$subject_id)

d_sub_wide <- d_sub_long |>
  group_by(subject_id, 
```

>>>>

```{r}
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(psych))
suppressPackageStartupMessages(source(here("helper","common.R")))
d_sub <- readRDS(here("cached_intermediates","1_d_sub.Rds"))
```

@@

```{r, fig.width=9, fig.height=9}
GGally::ggpairs(select(d_sub, age, log_age, rt, log_rt, 
                       short_window_accuracy, long_window_accuracy, 
                       short_window_elogit, long_window_elogit, prod, comp), 
                progress = FALSE, lower = list(continuous = GGally::wrap("points", alpha = 0.03)))

@@

```{r, fig.width=9, fig.height=9}
GGally::ggpairs(select(d_sub, log_age, rt,  
                       long_window_accuracy, short_window_accuracy, prod, comp), 
                progress = FALSE, lower = list(continuous = GGally::wrap("points", alpha = 0.03)))

@@

```{r}
m_rt <- mean(d_sub$rt, na.rm=TRUE)
sd_rt <- sd(d_sub$rt, na.rm=TRUE)
d_sub$factor_rt <- cut(d_sub$rt, 
                           breaks = quantile(d_sub$rt, 
                                             probs = c(0, .25, .5, .75, 1), 
                                             na.rm= TRUE), 
                           labels = c("0-25", "25-50", "50-75", "75-1"))

ggplot(filter(d_sub, !is.na(rt), !is.na(factor_rt)),
       aes(x = age, 
           y = long_window_accuracy, 
           col = factor_rt)) +
  geom_point(alpha = .3) +
  geom_smooth() +
  scale_x_log10() + 
  geom_hline(yintercept = .5, lty = 2)+
  viridis::scale_color_viridis(discrete = TRUE, 
                               name = "RT quantile") + 
  xlab("Age (log months)") + 
  ylab("Accuracy (long window)") + 
  theme(legend.position = "bottom")

@@

d_sub$age_group <- cut(d_sub$log_age, 
                           breaks = log(c(12,24,36,48,60)), 
                           labels = c("12-24","24-36","36-48","48-60"), 
                           include_lowest = TRUE)

ggplot(filter(d_sub, !is.na(age_group)), 
       aes(x = rt, 
           y = long_window_accuracy, 
           col = age_group)) +
  geom_point(alpha = .3) +
  geom_smooth() +
  geom_hline(yintercept = .5, lty = 2)+
  viridis::scale_color_viridis(discrete = TRUE, 
                               name = "age group")
```

@@

ggplot(filter(d_sub, 
              !is.na(rt), !is.na(factor_rt), 
              dataset_name %in% datasets_with_age_variation$dataset_name),
       aes(x = age, 
           y = long_window_accuracy, 
           col = factor_rt)) +
  geom_point(alpha = .3) +
  geom_smooth(se= FALSE, method = "lm") +
  scale_x_log10() + 
  geom_hline(yintercept = .5, lty = 2)+
  viridis::scale_color_viridis(discrete = TRUE, 
                               name = "RT quantile") +
  facet_wrap(~dataset_name) + 
  xlab("Age (log months)") + 
  ylab("Accuracy (long window)")

@@

```{r}
d_sub_mat <- d_sub |>
  ungroup() |>
  select(log_age, rt,  long_window_accuracy, prod)

d_sub_mat <- d_sub_mat |>
  filter(complete.cases(d_sub_mat)) |>
  as.matrix()

acc_rt_prc <- prcomp(x = d_sub_mat, scale = TRUE)

@@

```{r}
mod1 <- fa(d_sub_mat, 1)
mod2 <- fa(d_sub_mat, 2)
```

@@

```{r}
fa.parallel(d_sub_mat)
```

@@

```{r}
d_sub_mat_comp <- d_sub |>
  ungroup() |>
  select(log_age, rt,  long_window_accuracy, comp)

d_sub_mat_comp <- d_sub_mat_comp |>
  filter(complete.cases(d_sub_mat_comp)) |>
  as.matrix()

acc_rt_prc_comp <- prcomp(x = d_sub_mat_comp, scale = TRUE)

@@

```{r}
mod1_comp <- fa(d_sub_mat_comp, 1)
mod2_comp <- fa(d_sub_mat_comp, 2)
```

```{r}
fa.parallel(d_sub_mat_comp)
```

@@

```{r}
longitudinal <- d_sub |>
  group_by(dataset_name, subject_id) |>
  count() |>
  filter(n > 1)

d_sub_long <- d_sub |>
  filter(subject_id %in% longitudinal$subject_id, 
                     !is.na(prod)) |>
  group_by(subject_id) |>
  arrange(administration_id) |>
  mutate(admin_num = 1:n(), 
         delta_t = c(0,diff(age)))

ggplot(filter(d_sub_long, admin_num > 1), 
       aes(x = delta_t)) +
  geom_histogram()
```
